```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(here) #easy file paths
library(tidyverse) #data wrangling
library(tidymodels) #modeling
library(tidyclust)
library(patchwork)

set.seed(42) #set random seed
```

## Data Cleaning

Today, we are going to use the top 3 habitats and take a random sample of 500, then select trophic level and length for clustering

```{r}
# clean data
fish_clean = read_csv(here("discussion", "data", "w2-fish-clean.csv")) %>% 
  mutate(habitat = as.factor(habitat)) %>% 
  
  # filtered to 3 most different habitat to improve results
  filter(habitat %in% c("reef-associated", "demersal", "benthopelagic")) %>% 
  sample_n(500)

# we're clustering based on trophic level & length so we're working in 2D space
fish_clust <- fish_clean %>% 
  select(trophic_level_impute, length_impute) 

# view class of cluster columns
lapply(fish_clust, class)

```

## Clustering

```{r}
folds <- vfold_cv(fish_clust, v=5)
```

### KMeans

```{r}
kmeans_spec <- k_means(num_clusters = tune()) %>% 
  set_engine("stats") #stats selects base R kmeans function
```

```{r}
recipe <- recipe(~., data = fish_clust) %>% 
  step_normalize(all_predictors())

kmean_workflow <- workflow(recipe, kmeans_spec)
```

```{r}
# creating a grid of 1-10 cluster centers that are used
clust_num_grid <- grid_regular(num_clusters(),
                               levels = 10) 
```

```{r}
res <- tune_cluster(
  kmean_workflow,
  resamples = folds,
  grid = clust_num_grid,
  control = control_grid(save_pred = TRUE, extract = identity),
  metrics = cluster_metric_set(sse_within_total, sse_total, sse_ratio)
)
```
Inputs are similar to other tunes, inputing wf, grid, and folds. The metrics are some of the most common for kmeans. We want the `sse_ratio` to be as small as possible, as this signifies that values between points are small. `sse_ratio` is calculated using `sse_within_total` and `sse_total`. 

```{r}
res_metrics <- res %>% collect_metrics()
res_metrics
```
Produces mean values of the metrics we set. For example, in then `num_clusters` is 1, the mean ratio (mean column) is 1 since all points are within a single cluster.

#### Comparing values of k

```{r}
res_metrics %>%
  filter(.metric == "sse_ratio") %>%
  ggplot(aes(x = num_clusters, y = mean)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  ylab("mean WSS/TSS ratio, over 5 folds") +
  xlab("Number of clusters") +
  scale_x_continuous(breaks = 1:10)
```
The ratios printed above using `collect_metrics` are the same values used here to find the "elbow". We'll be comparing k=4 and k=5, since there is no way to use `select_best` to pull out the true best value.

```{r}
# create specifications with 4 clusters
kmeans_spec1 <- k_means(num_clusters = 4) %>% 
  set_engine("stats")

# create workflow using new specs and same recipe
kmeans_workflow1 <- workflow(recipe, kmeans_spec1)
```

```{r}
kmeans_fit1 <- kmeans_workflow1 %>% 
  fit(data = fish_clust)
kmeans_fit1
  
```
`between_ss` is the within cluster variance, and you want this to be small. This means the values within the clusters are very similar. You would also want the `total_ss` to be larger. This indicates high variation between different clusters.

```{r}
# create specifications with 5 clusters
kmeans_spec2 <- k_means(num_clusters = 5) %>% 
  set_engine("stats")

# create workflow using new specs and same recipe
kmeans_workflow2 <- workflow(recipe, kmeans_spec2)
```

```{r}
kmeans_fit2 <- kmeans_workflow2 %>% 
  fit(data = fish_clust)
kmeans_fit2
```
The ratio of within/total is slightly higher. This means the model performs slightly better than when k=5, but this should not be used as the only metric for choosing k.

#### Comparing silhouette
```{r}
kmeans_fit1 %>% 
  silhouette_avg(fish_clust)
```

```{r}
kmeans_fit2 %>% 
  silhouette_avg(fish_clust)
```
The `silhouette_avg` checks validatiy and accuracy within the data. The higher `.estimate` of the silhouette is the better performing model. 

#### Plotting clusters
```{r}
# isolate cluster assignments for points
clusters <- kmeans_fit1 %>% 
  extract_cluster_assignment()

# append cluster column onto original df
fish_clean$cluster = clusters$.cluster
```

```{r}
ggplot(fish_clean) +
  geom_point(aes(trophic_level_impute, length_impute,
                 color = cluster, shape = habitat)) +
  theme_minimal() +
  scale_color_manual(values = c("maroon", "dodgerblue", "yellowgreen",
                                "mediumorchid"))
```
Added the `shape = cluster` to see if the habitat types coincide with the clusters. Overall the model did pretty poorly, likely because there is no clear clustering in the data provided.

### Heirarchical Clustering

```{r}
# apply hierarchical clustering
hc_spec <- hier_clust(
  linkage_method = "average"
)
```
`linkage_method` defines how the agglomaration is done. Hasn't been gone over in lecture, so we won't go over it now.

```{r}
# fit model same way, no recipe bc you don't need to normalize for heirarchical clustering
hc_fit <- hc_spec %>% 
  fit(~., data = fish_clean)
```

```{r}
# plot fit data
hc_fit$fit %>% plot
```
Very ugly. Fish 380 is very different, it separates super early.
