---
title: "Lab 3 Demo"
author: "Briana Baraas"
date: "2023-01-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rsample)
library(skimr) #exploratory package for dataset summaries
library(glmnet)
```

## Data Wrangling and Exploration

```{r data}
#load and inspect the data
dat <- AmesHousing::make_ames()

```

##Train a model

```{r intial_split}
# Data splitting with {rsample} 
set.seed(123) #set a seed for reproducibility

# split data (default prop is 3/4)
split <- initial_split(dat)

# isolate training and test data
ames_train <- training(split)
ames_test  <- testing(split)

```

```{r model_data}
#Create training feature matrices using model.matrix() (auto encoding of categorical variables)

# creating the model matrix puts data in a form that the caret uses
  # outcome variable ~ predictor, use . to select all predictors 
  # -1 removes col for intercept variable (not needed, only has value 1)
X <- model.matrix(Sale_Price ~ ., ames_train)[,-1]  

# transform y with log() transformation
Y <- log(ames_train$Sale_Price)

# View data, Sale_Price is heavily right skewed & can be log transformed
skim(dat)
```

```{r glmnet}
#fit a ridge model, passing X,Y,alpha to glmnet()
ridge1 <- glmnet(x = X, y = Y,
                alpha = 0) #set to 0 for ridge model

#plot() the glmnet model object
plot(ridge1, 
     xvar = 'lambda') #variable to plot
```

**Interpretation:** Plot shows that lambda is shrinking coefficients for the predictor variables.

```{r}
# lambdas applied to penalty parameter.  Examine the first few
ridge1$lambda %>% head()

# small lambda results in large coefficients
coef(ridge1)[c("Latitude", "Overall_QualVery_Excellent"), 100]

# what about for small coefficients?
coef(ridge1)[c("Latitude", "Overall_QualVery_Excellent"), 1]

```

**Interpretation:** The 100 and 1 represent the column that you are indexing (?). When lambda is larger (100) we see that `Latitude` has a smaller lambda, meaning it has a more significant impact on the model than `Overall_QualVery_Excellent`.

How much improvement to our loss function as lambda changes?

##Tuning
We'll be using cross-validation, and re-sampling the data and running the model on a portion of the data and averaging the error for different folds. The default is a 10-fold sample. 
```{r cv.glmnet}
# Apply CV ridge regression to Ames data.  Same arguments as before to glmnet()
ridge2 <- cv.glmnet(x = X, y = Y,
                   alpha = 0) #ridge

# Apply CV lasso regression to Ames data
lasso <- cv.glmnet(x = X, y = Y,
                   alpha = 1) #laso
  
# plot results
par(mfrow = c(1, 2))
plot(ridge2, main = "Ridge penalty\n\n")
plot(lasso, main = "Lasso penalty\n\n")
```
**Interpretation:** As lambda is increasing, the mean-squared error changes. For a good model, the MSE should be a lower value. Over the set of lambdas we explored, it looks like constraining the coefficients too much (higher lambda) decreases the model's performance. The first dotted line indicates the value of lambda that returns the lowest MSE. The second line gives you the most parsimonious model (fewer variables) that's within 1 standard error of the first line. 

10-fold CV MSE for a ridge and lasso model. What's the "rule of 1 standard deviation"?

In both models we see a slight improvement in the MSE as our penalty log(λ) gets larger, suggesting that a regular OLS model likely overfits the training data. But as we constrain it further (i.e., continue to increase the penalty), our MSE starts to increase.

Let's examine the important parameter values apparent in the plots.

```{r}
# ============= Ridge model =============
# minimum MSE
min(ridge2$cvm)

# lambda for this min MSE
ridge2$lambda.min

# 1-SE rule
ridge2$cvm[ridge2$lambda == ridge2$lambda.1se]

# lambda for this MSE
ridge2$lambda.1se

# ============ Lasso model ============

# minimum MSE
min(lasso$cvm)

# lambda for this min MSE
lasso$lambda

# 1-SE rule
lasso$cvm[lasso$lambda == lasso$lambda.1se]

# lambda for this MSE
lasso$lambda.1se

# No. of coef | 1-SE MSE (number of coefficients/predictors is 169)
lasso$nzero[lasso$lambda == lasso$lambda.min]
```

```{r}
# Ridge model
ridge_min <- glmnet()

# Lasso model
lasso_min


par(mfrow = c(1, 2))
# plot ridge model
plot(ridge_min, xvar = "lambda", main = "Ridge penalty\n\n")
abline(v = log(ridge$lambda.min), col = "red", lty = "dashed")
abline(v = log(ridge$lambda.1se), col = "blue", lty = "dashed")

# plot lasso model
plot(lasso_min, xvar = "lambda", main = "Lasso penalty\n\n")
abline(v = log(lasso$lambda.min), col = "red", lty = "dashed")
abline(v = log(lasso$lambda.1se), col = "blue", lty = "dashed")
```
