---
title: "Lab4_Demo"
author: "Mateo Robbins"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # for data wrangling, and plotting
library(rsample)   # for data splitting
library(recipes)   # data pre-processing
library(caret)     # for logistic regression modeling
```
Let's explore how employee income and overtime hours worked affect likelihood of employee attrition.  Any predictions?

__Prediction:__ Attrition is when 

```{r}
#
data("attrition", package = "modeldata")

df <- attrition %>% mutate_if(is.ordered, factor, ordered = FALSE)

# Create training (70%) and test (30%) sets for the 
# rsample::attrition data (old, don't need this code)
set.seed(123)  # for reproducibility 
churn_split <- initial_split(df, prop = 0.7)
churn_train <- training(churn_split)
churn_test  <- testing(churn_split)
```

Let's set up set up our recipes now for preprocessing. 
```{r recipe}
#specify and prep recipe
churn_rec <- recipe(Attrition~., data = churn_train) %>% 
  
  # convert attrition into 0/1 binary variable 
  step_integer(Attrition, zero_based = TRUE) %>%
  
  prep(churn_train)

#bake recipe with training data 
churn_baked_train <- bake(churn_rec, new_data = churn_train)

# bake will convert attrition to 0/1, as defined by the recipie
unique(churn_baked_train$Attrition)
```


```{r specify_models_glm}
#MonthlyIncome
model_inc <- glm(data = churn_baked_train,
                 # input predictor variable (income)
                 Attrition~ MonthlyIncome, 
                 # specify type of model
                 family = 'binomial')

#OverTime
model_time <- glm(data = churn_baked_train,
                  Attrition~OverTime,
                  family = 'binomial')
```


```{r tidy_model_objs}
# convert both models to tidy form
tidy(model_inc)
tidy(model_time)
```

```{r exp_coefs}
#exponentiate the coefficients from model objects for interpretation. Gives us changes in odds of attrition

# exponentiate model coefficients for income
exp(coef(model_inc))
```
__Given Interpretation:__ For each dollar an employee makes the likelihood of them leaving decreases. We know it's decreasing since 0.999 is a decimal, and multiplying by this value decreases

__My Interpretation:__ For each additional dollar an employee makes, the _odds_ of them leaving decreases by 0.999. 

```{r}
exp(coef(model_time))
```
__Interpretation:__ The _odds_ of an employee leaving increases by 3.48 times if they worked overtime. 

```{r recode_attrition_test}
# run model w/data it hasn't seen (test data)
# start by applying recipie to test data
churn_baked_test <- bake(churn_rec,
                         new_data = churn_test)
```

```{r plot_income_attrition}
# plot income as single predictor
ggplot(churn_baked_test, aes(x = MonthlyIncome,
                             y = Attrition)) +
  geom_point() +
  stat_smooth(method = 'glm', se = TRUE, 
              method.args = list(family = 'binomial'))
```

We can add more predictors, creating a multiple logistic regression model

```{r mult_log_regression}
# train multi-variable model
model_both <- glm(Attrition ~ MonthlyIncome + OverTime,
                  data = churn_train,
                  family = 'binomial')

# tidy multi-variate model
tidy(model_both)

# plot
ggplot(churn_baked_test, aes(x = MonthlyIncome, 
                             y = Attrition,
                             color = OverTime)) +
  geom_point(alpha = 0.3) +
  stat_smooth(method = 'glm', se = FALSE,
              method.args = list(family = 'binomial'))
```

